---
title: "p8105_hw2_yz4436"
author: "Yuanhao Zhang"
date: "2022-10-04"
output: github_document
---

```{r load_libraries}
library(tidyverse)
library(readxl)
```

### Problem 1

Import and clean date from `NYC_Transit_Subway_Entrance_And_Exit_Data.csv` and changed route 8-11 to character. Then change variable entry to logical variable and select something that will be used later.
```{r}
trans_ent = 
  read_csv(
    "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

As it stands, these data are not "tidy": route number should be a variable, as should route. That is, to obtain a tidy dataset we would need to convert `route` variables from wide to long format. This will be useful when focusing on specific routes, but may not be necessary when considering questions that focus on station-level variables. 

The following code chunk selects station name and line, and then uses `distinct()` to obtain all unique combinations. As a result, the number of rows in this dataset is the number of unique stations.

```{r}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```

The next code chunk is similar, but filters according to ADA compliance as an initial step. This produces a dataframe in which the number of rows is the number of ADA compliant stations. 

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

To compute the proportion of station entrances / exits without vending allow entrance, we first exclude station entrances that do not allow vending. Then, we focus on the `entry` variable -- this logical, so taking the mean will produce the desired proportion (recall that R will coerce logical to numeric in cases like this).

```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

Lastly, we write a code chunk to identify stations that serve the A train, and to assess how many of these are ADA compliant. As a first step, we tidy the data as alluded to previously; that is, we convert `route` from wide to long format. After this step, we can use tools from previous parts of the question (filtering to focus on the A train, and on ADA compliance; selecting and using `distinct` to obtain dataframes with the required stations in rows).

```{r}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```


### Problem 2

Read and clean the Mr. Trash Wheel sheet with specifying the sheet in the Excel file and to omit non-data entries using arguments in read_excel, using reasonable variables, omitting rows that do not include dumpster-specific data, and rounding the number of sports balls to the nearest integer and converts the result to an integer variable. Then I create a new variable to indicate this is Mr. Trash Wheel and locate it in the first column.
```{r}
mr_trashwheel_df =
  read_excel(
    "Data/Trash Wheel Collection Data.xlsx", 
    sheet = "Mr. Trash Wheel",
    range = "A2:N549") %>%
  janitor::clean_names() %>%
  mutate(
    sports_balls = as.integer(round(sports_balls, 0)),
    year = as.numeric(year),
    Trash_Wheel_Type = "Mr." 
  ) %>%
  relocate(Trash_Wheel_Type)
```

Import, clean, and organize the data for Professor Trash Wheel, and then I create a new variable to indicate this is Professor Trash Wheel and locate it in the first column. Finally, combining this with the Mr. Trash Wheel dataset to produce a single tidy dataset.
```{r}
professor_trashwheel_df =
  read_excel(
    "Data/Trash Wheel Collection Data.xlsx", 
    sheet = "Professor Trash Wheel",
    range = "A2:M96") %>%
  janitor::clean_names() %>%
  mutate(Trash_Wheel_Type = "Professor") %>%
  relocate(Trash_Wheel_Type)

combine_tidy = bind_rows(mr_trashwheel_df, professor_trashwheel_df)
```
